---
title: "Hand-on Exercise 1"
subtitle: "Exploring ggplot2 methods"
author: "Fangxian"
---

# Getting Started

## Install and launching R packages

The code chunk below uses p_load() of pacman pacakge to check if tidyverse packages are installed in the comupter. If they are, then they will be launched into R.

```{r}
pacman::p_load(tidyverse)

```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# Essential Grammatical Elements in ggplot2: geom

## Plotting a simple bar chart

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar()
```

### Modifying a geometric object by changing geom()

```{r}
ggplot(data=exam_data,
       aes(x= MATHS)) +
  geom_histogram(bins = 20,
                 color="black",
                 fill= "light blue")
```

### Modifying a geometric object by changing aes()

```{r}
ggplot(data=exam_data,
       aes(x= MATHS,
           fill = GENDER)) +
  geom_histogram(bins=20,
                 color="grey30")
```

## Geometric Objects: geom_density()

This is a useful alternative to the histogram for continuous data that comes from an underlying smooth distribution.

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_density()        
```

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS, 
           colour = GENDER)) +
  geom_density()
```

## Geometric Objects: geom_boxplot

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS,       
           x= GENDER)) +    
  geom_boxplot()            
```

Adding in notches

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot(notch=TRUE)
```

## Geometric Objects: geom_violin

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_violin()
```

## Geometric Objects: geom_point()

This is useful for creating scatterplot.

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point()  
```

## *geom* objects can be combined

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot() +                    
  geom_point(position="jitter", 
             size = 0.5)
```

# Essential Grammatical Elements in ggplot2: stat

## Working with stat - the stat_summary() method

below code showing mean in red dots.

```{r}
ggplot(data = exam_data,
       aes(y = MATHS, x = GENDER)) +
  geom_boxplot() +
  stat_summary(geom = "point",       
               fun.y="mean",         
               colour ="red",        
               size=4) 
```

## Working with stat - the geom() method

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  geom_point(stat="summary",        
             fun.y="mean",           
             colour ="red",          
             size=4)  
```

## Adding a best fit curve on a scatterplot

using geom_smooth(size=0.5)

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(size=0.5)
```

The default smoothing method (default is loess) can be overridden as shown below.

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5)
```

# Essential Grammatical Elements in ggplot2: Facets

## Working with facet_wrap()
