---
title: "In-Class Exercise 4"
execute: 
  eval: true
  echo: true
  warning: false
author: "Fangxian"
date: "6 May 2023"
date-modified: "`r Sys.Date()`"
---

### Getting Started

```{r}
pacman::p_load(rstatix, gt, patchwork, tidyverse)
```

```{r}
#| echo: true
exam_data <- read_csv("data/Exam_data.csv")
```

### Visualising Normal Distribution

If the data is normally distributed, the points in a Q-Q plot will lie on a straight diagonal line. Conversely, if the points deviate significantly from the straight diagonal line, then it's likes likely that the data is normally distributed.

::: panel-tabset
## The plot

```{r}
#| echo: false
#| fig-width: 4
#| fig-height: 4
ggplot(exam_data,
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()
```

## The code chunk

```{r}
#| eval: false
#| fig-width: 4
#| fig-height: 4
ggplot(exam_data,
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()
```
:::

::: callout - Note Note: We can see that the points deviate significantly from the straight diagonal line. This is a clear indication that the set of data is not normally distributed. :::

### Combining satistical graph and analysis table

Need to install webshot

::: panel-tabset

## The plot

```{r}
#| echo: false
qq <- ggplot(exam_data,
       aes(sample=ENGLISH)) +
    stat_qq() +
    stat_qq_line()

sw_t <- exam_data %>%
  shapiro_test(ENGLISH) %>%
  gt()

#converting the image into a temp file
tmp <- tempfile(fileext = '.png')
gtsave(sw_t,tmp)
table_png <- png::readPNG(tmp, native = TRUE)

qq + table_png
```

## The code chunk

```{r}
#| eval: false
qq <- ggplot(exam_data,
       aes(sample=ENGLISH)) +
    stat_qq() +
    stat_qq_line()

sw_t <- exam_data %>%
  shapiro_test(ENGLISH) %>%
  gt()

#converting the image into a temp file
tmp <- tempfile(fileext = '.png')
gtsave(sw_t,tmp)
table_png <- png::readPNG(tmp, native = TRUE)

qq + table_png
```
:::

